---
title: "project"
author: "RawanGalal"
date: "3/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r imports}
library(tidyverse)
library(stringr)
library(lubridate)
library(validate)
library(janitor)

```


```{r}
options(scipen = 50)
spec_csv("Data.csv")

```

```{r read_file}
data = read_csv("Data.csv")
data
```

# Level 0: Format and file structure
#### this is transactional data, in the long format

```{r unique_rows}
data %>% distinct()

```

#### :87,396 rows are distinct
#### check duplicates. is there a pattern? 

```{r duplicates_show}
data %>% 
#distinct(hotel, is_canceled, lead_time, arrival_date_year, arrival_date_month,  arrival_date_week_number, #arrival_date_day_of_month, stays_in_week_nights, stays_in_weekend_nights, adults, children, babies, meal, country, #market_segment, distribution_channel, is_repeated_guest) %>%
 group_by(hotel, is_canceled, lead_time, arrival_date_year, arrival_date_month, arrival_date_week_number, arrival_date_day_of_month, stays_in_week_nights, stays_in_weekend_nights, adults, children, babies, meal, country, market_segment, distribution_channel, is_repeated_guest, previous_cancellations,previous_bookings_not_canceled,  reserved_room_type, assigned_room_type, booking_changes, deposit_type, agent, company, days_in_waiting_list, customer_type, adr, required_car_parking_spaces, total_of_special_requests, reservation_status, reservation_status_date) %>%
  filter(n() >1) %>%
  ungroup() %>%
  arrange(hotel, is_canceled, lead_time, arrival_date_year, arrival_date_month, arrival_date_week_number, arrival_date_day_of_month, stays_in_week_nights, stays_in_weekend_nights, adults, children, babies, meal, country, market_segment, distribution_channel, is_repeated_guest, previous_cancellations,previous_bookings_not_canceled,  reserved_room_type, assigned_room_type, booking_changes, deposit_type, agent, company, days_in_waiting_list, customer_type, adr, required_car_parking_spaces, total_of_special_requests, reservation_status, reservation_status_date)
  
 # data[!duplicated(data)]
```


#### conclusion; there are no duplicates; two people can book at the same hotel, through the same agent, arrvie at the same day...


# ################################################
# Level 1: cells
## a) aggregate summary statistics
```{r stats_numeric}
data %>%
  select(where(is.numeric)) %>%
  summary()

```

#### worth investigation: 1. max. in lead_time; is it possible to book in a hotel before arriving with 2 years?
#### 2. adults; 55 ????? It seems like booking the entire hotel not a room!
#### 3. NA's in children>> see distribution
#### 4. 10 babies????
#### outliers: stays_in_weekend_nights,  stays_in_week_nights, children, previous_cancellations, 
#### previous_bookings_not_canceled, booking_changes, days_in_waiting_list(more than 1 year in waiting list??), adr??
#### required_car_parking_spaces?  

### lead_time
```{r lead_time_outliers}
data %>% summarise(quantile(lead_time, 0.999999))


```

#### 99% of the data is less than 444
#### let's dive deeper 
```{r}
data %>% 
  filter( lead_time > 733)
```

```{r adults_outliers}
data %>% summarise(quantile(adults, .9999))


```

```{r }
data %>% 
  filter(adults> 10) 
```
```{r na_children}
data %>%
  filter(is.na(children))
```
#### since there are only 4 rows, with no obvious pattern, I decided to keep these rows since the number of children won't affect the analysis (untill now)


```{r babies_outliers}
data %>%summarise(quantile(babies, .9999))

```

```{r}
data %>%
  filter(babies >2)

```
```{r stays_in_weekend_nights_Outliers}
data %>%
  summarise(quantile(stays_in_weekend_nights, 0.99))

```
```{r}
data %>%
  filter(stays_in_weekend_nights > 12)

```
#### this outlier doesn't indicate a data error; a person can stay any number of weekend nights. However, it does #### tell us that 99% of guests stayed only 4 weekend nights

```{r children_outliers}
data %>%summarise(quantile(children, .9999, na.rm = T))

```
```{r}
data %>%
  filter( children > 3)

```

#### only 1 row. Most propably, it's a data entry error. **But let's make sure from the business first.**

```{r}

data %>% 
  summarise(quantile(previous_cancellations, .999))
```

```{r}
data %>% filter(previous_cancellations >14) %>% distinct()

```

```{r}
data %>% filter(previous_cancellations >14, is_repeated_guest ==0)

```

#### This is worth highliting. Having a high cancellation rate represents a threat to the business. 
#### **FURTHER INVESTIGATIONS** need to be done to reveal any hidden patterns. For instance, could agents; 208, 252, 96, 1 relate to #### this event? country; PRT? Room type=A? **they are not repeated guests (except for 1)**
#### most of them booked  ~ 200 days in advance.

#### the same thinking approach applies to previous_bookings_not_canceled; there's an outlier of 72 which deserves more attention; were these bookings made throuhg a certain period of time? could I reveal patterns that helps me maximize the profit?

```{r booking_changes_outliers}
data %>%summarise(quantile(booking_changes, .9999))
data
```


```{r}
data %>% filter(booking_changes >14) %>% distinct()

```

```{r  days_in_waiting_list_Outliers}
data %>% summarise(quantile ( days_in_waiting_list, 0.99))

```
```{r}
data %>% 
  filter( days_in_waiting_list >75)

```

```{r  adr_Outliers}
data %>% summarise(quantile ( adr, 0.9999))

```

```{r}
data %>% 
  filter( adr > 388)

```

### b) frequency table of categorical columns
```{r}
tabyl(data$hotel)
tabyl(data$arrival_date_month)
tabyl(data$meal)
tabyl(data$country)
tabyl(data$market_segment)
tabyl(data$distribution_channel)
tabyl(data$reserved_room_type)
tabyl(data$assigned_room_type)
tabyl(data$deposit_type)
tabyl(data$agent) %>% filter(n !=2)
tabyl(data$company) %>% filter(n !=1)
tabyl(data$customer_type)
tabyl(data$reservation_status)
```
### - agent column
#### there are 16340 nulls in the agent column. >> 13.7%
```{r agent_nulls}

data %>% group_by(agent) %>%
  filter(n() ==16340)
```
#### 36% of the market segment with null agent are direct 
#### 69.7% of company with null agent are alsoo null 
#### 71.7% of customers with null agent are transient
#### 75.3% of reservation_status with null agents are check-out


### - company column
#### 94% of the copmany column is null> **drop**

### - reservation_status column
#### 36% of reservations are cancelled!! this is a high rate and needs **further investigations**


### Intuitively, behaviors will differ in a city hotel rather than a resort hotel.
### so I'll separate them for future analysis
```{r}
data_city <- data %>% filter(hotel== "City Hotel")
data_resort <- data %>% filter(hotel == "Resort Hotel")
tabyl(data_city$arrival_date_month)
tabyl(data_resort$arrival_date_month)
#data_resort 
```


# #####################################################
## rationality of each column
```{r}
data %>% mutate(year_val =  str_split(reservation_status_date,  "-")[0])


```

# ############################################################################
# b) cross-column validations
## 1b- date vs year/month/day
**come back to it **


## 1c - market_segment vs distribution channel


#### arrival_date_day_of_month; validate; feb; 28?
#### arrive between the 1st of July of 2015 and the 31st of August 2017

#### visualizations 