---
title: "project_analysis"
author: "RawanGalal"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r imports}
library(tidyverse)
library(stringr)
library(lubridate)
library(ggplot2)
library(janitor)
library(plotly)

```

```{r read_file}
data = read_csv("Data.csv")
data
```

```{r}
data_city <- data %>% filter(hotel== "City Hotel")
data_resort <- data %>% filter(hotel == "Resort Hotel")
tabyl(data_city$arrival_date_month)
tabyl(data_resort$arrival_date_month)
#data_resort 
```

# Q1: is there a certain market segment that increases no. of transactions? 
```{r market_segment_vs_trans}
data_resort %>% tabyl(market_segment)

```
## Understanding market segments in hotels:
### 2. Corporate – These are the guests that stay at the hotel at discounted company rates. The discounted rates are available on a contract basis to corporate clients who bring in frequent business.
### 3. Direct –This segment consists of direct bookings with the hotel done over phone calls or emails.
### 5. Online TA: online third-party travel agencies that allow people to make reservations for various hotels through their platform. For example, Trivago, TripAdvisor,

### 44% of market segment in resort hotels are online TA, while only .5% are complementary
### insight: maintain the online TA>> turn them into loyal customers, and work on increasing the comp. cusstomers
```{r}
g <- ggplot(data, aes(market_segment))
g + geom_bar()
data
```
### compare demand in peak months(July and august, which correspond to 33% of the total demand ) for each segment. does it increase over years?


```{r}
#tabyl(data_resort$arrival_date_month)
# cond: market segment
#value: trxs
# x: year buckets
 data_resort %>% filter(arrival_date_month == "July" | arrival_date_month == "August") %>%
 group_by(arrival_date_year, market_segment) %>%
 summarise(number_of_transactions_in_7_and_8 = n() , market_segment = market_segment) %>%
 ggplot( aes(factor(arrival_date_year), number_of_transactions_in_7_and_8 , fill = market_segment)) + 
 geom_bar(stat="identity", position = "dodge") + 
 scale_fill_brewer(palette = "Set1")

```

#### *insights* : 1. Online TA reached is increasing over years in peak months > recommendations: maintain their loyalty > points, discounts, through TA ...
### 2. direct customers have a slight decrease over time 
### 3. corporate is consistently low>> partner with more companies or cancel current contracts?
### 4. offline TA is decreasing over time. a possible revenue threat? >> the decrease is small and doesn't worth investing in a marketing campaign; replace current offline marketing TA's?

```{r}


# data_resort %>%
# group_by(arrival_date_year, market_segment ) %>%
# summarise(number_of_transactions = n() , market_segment = market_segment, arrival_date_month = arrival_date_month) #%>%
#ggplot(aes(x = arrival_date_month, y = number_of_transactions)) + geom_line(aes(colour = factor(arrival_date_year)))
#
#
#
#
# data_resort %>% filter(arrival_date_year == 2017  & market_segment == "Online TA") %>%
# group_by(arrival_date_month) %>%
# summarise(number_of_transactions = n() ) %>%
# ggplot( aes(arrival_date_month, number_of_transactions)) +
#   geom_line()
 
 
 
#  data_resort %>% filter(arrival_date_year == 2016 & market_segment == "Online TA") %>%
# group_by(arrival_date_month) %>%
# summarise(number_of_transactions = n() ) %>%
#    plot_ly(x = ~arrival_date_month, y = ~ number_of_transactions, type = 'scatter', mode = 'lines')
# 
# 
# data_resort %>% filter(arrival_date_year == 2017 & market_segment == "Online TA") %>%
# group_by( arrival_date_month) %>%
# summarise(number_of_transactions_17 = n() ) %>%
#   
#  plot_ly(x = ~arrival_date_month, y = ~number_of_transactions_17, type = 'scatter', mode = 'lines')

 #+  plot_ly(x = ~arrival_date_month, y = ~ number_of_transactions, type = 'scatter', mode = 'lines')

   
 

```



```{r}
# cond: market segment
#value: trxs
# x: year buckets

#data_resort %>% filter(arrival_date_year == 2016) %>%
#group_by(arrival_date_month, market_segment) %>%
#summarise(number_of_transactions = n() , market_segment = market_segment) %>%
#ggplot( aes(factor(arrival_date_month), number_of_transactions, fill = market_segment)) + 
#geom_bar(stat="identity", position = "dodge") + 
#scale_fill_brewer(palette = "Set1")

```

### same analysis for city hotels
```{r}
tabyl(data_city$market_segment)

```
```{r}
#tabyl(data_resort$arrival_date_month)
# cond: market segment
#value: trxs
# x: year buckets
 data_city%>% filter(arrival_date_month == "July" | arrival_date_month == "August") %>%
 group_by(arrival_date_year, market_segment) %>%
 summarise(number_of_transactions_in_7_and_8 = n() , market_segment = market_segment) %>%
 ggplot( aes(factor(arrival_date_year), number_of_transactions_in_7_and_8 , fill = market_segment)) + 
 geom_bar(stat="identity", position = "dodge") + 
 scale_fill_brewer(palette = "Set1")

```

####Online TAs are again the dominant, with a continous increment> same reco. The pattern here deserves more attention; online TAs witnessed a rapid increase between 2015 and 2016 and a slight increase in 2017. why?
### was there an applied policy in 2016 that needs to be implemented? what changed?
#### direct is increaing, this propably means that more people started to hear about this hotel.
#### offline TA/TO >> same reco.
#### key factors that impact the booking: the channel through wich different market segments use to book.